---
title: "Study of automatic and manual transmission vs miles per galon"
author: "Salvador Castañeda"
date: "18 de julio de 2017"
output: pdf_document
---

## Introduction

This study aims to determine wether a manual or automatic transmission gives a best miles per galon perfomrance. 
The data set used for this assignment is `mtcars`. [here](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) you can find the full documentation.
Its important to remark that this data was collected in 1974 and therefore it is not applicable to the present due to a almost sure structural change.

## Summary
In the study we found that manual transmittion has a significant better performance over automatic transmission.
In fact, in average there is a 7.25 mpg increase in manual transmittion versus automatic though there are other important variables that impact the relationship.


## Getting into the matter 
We show the first rows of the dataset.

```{r, echo=FALSE}
library(ggplot2)
data(mtcars)
```


```{r}
head(mtcars)
```


For this study we´re going to ask the questions:

        1.-Is an automatic or manual transmission better for MPG?
        2.-Quantify the MPG difference between automatic and manual transmissions?

So lets get into the matter

Lets fit a model with am as a predictor of mpg.
```{r}
summary(lm(mpg~am,data = mtcars))
```

It is important to note that the intercept coefficient is the mean for automatic transmiissions and the am coefficient is the difference between manual and automatic transmission. Below we ilustrate this:
```{r}
by(mtcars$mpg,mtcars$am,mean)
```
For both values, the intercept and am coefficient are statistically significant (their p-values are < .005).
So the model fitted would have this form: $mpg = 17.147+7.245*manual$ so if the transmittion is manual it has on average a 7.245 increase on the mpg.

Now, the R-squared is 0.36 that suggests that mpg is not quite well estimated only by the am variable.
The following plot is aimed to show that transmission is not the only important variable affecting miles per galon.
Here we can see that as weight, horse power and number of cylinders increase, miles per galon decreases. 
Thus, either we have automatic or manual transmissors, there are other variables that imapct the profit of mpg.

```{r fig.width = 6, fig.height = 5,warning=FALSE}
i <- ggplot(mtcars, aes(x=hp,y=wt,colour=mpg,size=factor(cyl)))+geom_point()
i <- i+facet_grid(facet=.~factor(am))
i
```

Now we´re going to use the `bestglm` package and figure out the best combination of variables to fit our model.

```{r warning=FALSE}
library(bestglm)
mtcars <- mtcars[,c(2:11,1)]
bestglm(mtcars, family= gaussian)
summary(lm(mpg~wt+qsec+am,data=mtcars))
```

So now R-squared has grown to 85%.

```{r}
par(mfrow=c(2,2))
plot(lm(mpg~wt+qsec+am,data=mtcars))

```

##Finally, some exploratory analysis

First we make the classic boxplot because it makes visual the advantage of manual transmittions vs automatic ones.

```{r}

plot(factor(mtcars$am),mtcars$mpg,names=c("Automatic","manual"))
```

And to conclude, a plot using all the variables the `bestglm` found were the best combination. 

```{r}
ggplot(mtcars, aes(x=qsec,y=mpg,colour=factor(am),size=wt))+geom_point()
```



